On the specific recursive structure of iterated parametricity

### 2025-11-09 paper/paper.tex
paper: introduce more fulldeps

\newcommandx{\deps}[2][1,2]{\prim{deps}[][#1][#2][][]}
\newcommandx{\fulldeps}[2][1,2]{\prim{fulldeps}[][#1][][][]}
\newcommandx{\depscohs}[2][1,2]{\prim{depscohs}[][#1][#2][][]}
\newcommandx{\depscoht}[2][1,2]{\prim{depscoht}[][#1][#2][][]}
\newcommandx{\depstype}[2][1,2]{\prim{depstype}[][#1][#2][][]}
\newcommandx{\fulldepstype}[2][1,2]{\prim{fulldepstype}[][#1][][][]}
\newcommandx{\depscohstype}[2][1,2]{\prim{depscohstype}[][#1][#2][][]}
\newcommandx{\depscohttype}[2][1,2]{\prim{depscohttype}[][#1][#2][][]}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n][0,...,p]   &
    (\fulldeps[n-1])        & :      & \paintingtype[n][0,...,p](\framep[n][0,...,p](\fulldeps.\deps[n-1][0,...,p])) \\
    \painting[n][0,...,0-1] &
    \left(\begin{array}{l}
              \unitpoint \\
              E          \\
            \end{array}\right)  & \defeq & \unittype                                                                   \\
    \painting[n][0,...,p]   &
    (\fulldeps[n-1])                                                                                                 \\
                            & \defeq &
    \left(\painting[n][0,...,p-1]
      (\fulldeps[n-1]),\,
    \painting[n][p]
    (\fulldeps[n-1])\right)                                                                                          \\
  \end{array}
\end{equation*}

@@

\begin{align*}
  \depscohstype[n][0,..,p-1] \defeq
  \left(\begin{array}{l}
            \deps[n][0,...,n] :\depstype[n][0,...,n]        \\
            \restrp[n][0,...p-1]: \restrptype[n][0,...,p-1] \\
            \cohf[n][0,...,p-1]: \cohftype[n][0,...,p-1]    \\
          \end{array}\right)
\end{align*}
so that the type of $\restrf[n][0,...,p]$ can be rephrased as $\Pi \depscohs[n][0,...,p-1]:\depscohstype[n][0,..,p-1]. \restrftype[n][0,...,p]      \left(\begin{array}{l}
      \depscohs.\deps.\framep[n][0,...,p]   \\
      \depscohs.\deps.\painting[n][0,...,p] \\
    \end{array}\right)$, and similarly for $\restrltype$ and $\restrl$. Later, we shall write $\depscohs.\deps$, $\depscohs.\restrp$, and $\depscohs.\cohf$ for the corresponding projections of an element $\depscohs$ in $\depscohstype$. By abuse of notation, we also write $\unitpoint$ for the only inhabitant of $\depscohstype[n][0,..,0-1]$.

@@

\begin{align*}
  \overline{\deps}(\depscohs:\depscohstype) : \depstype \\
  \overline{\deps}(\depscohs) \defeq
  \left(\begin{array}{l}
            \framep[n][p](\depscohs.\deps[n][0,...,p])   \\
            \painting[n][p](\depscohs.\deps[n][0,...,n]) \\
            \restrf[n][p](\depscohs[n][0,...,p])         \\
          \end{array}\right)    \\
\end{align*}

Armed with $\restrf$ and $\cohftype$, it is now possible to define
$\restrp[n][p]$ by recursion on $n-p$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrp[n][p]                                      &
    \left(\begin{array}{l}
              \depscohs[n][0,...,n]:\depscohstype[n][0,...,n] \\
              E:\framep[n-1][n-1]\rightarrow \U               \\
              E':\framep[n][n]\rightarrow \U                  \\
            \end{array}\right) & :      &
    \left(
    \begin{array}{l}
        \Pi q\leq n. \Pi \epsilon. \Pi d:\framep[n][p]
        \left(\begin{array}{l}
                  \overline{\deps}(\depscohs)    \\
                  E':\framep[n][n]\rightarrow \U \\
                \end{array}\right).    \\
        \painting[n][p]
        \left(\begin{array}{l}
                  \overline{\deps}(\depscohs)    \\
                  E':\framep[n][n]\rightarrow \U \\
                \end{array}\right)(d) \rightarrow \\
        \painting[n-1][p]
        \left(\begin{array}{l}
                  \depscohs[n][0,...,n].\deps       \\
                  E:\framep[n-1][n-1]\rightarrow \U \\
                \end{array}\right)(\restrf[n][p](\depscohs[n][0,...,n].deps)(d))\end{array}\right)                                                  \\
    \restrp[n][n]                                      &
    \left(\begin{array}{l}
              \depscohs[n][0,...,n] \\
              E                     \\
              E'                    \\
            \end{array}\right)                           & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c). l\epsilon \\
    \restrp[n][p](p<n)                                 &
    \left(\begin{array}{l}
              \depscohs[n][0,...,n] \\
              E                     \\
              E'                    \\
            \end{array}\right)                           & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c).
    \left(\begin{array}{l}
              \restrl[n][p][q][\epsilon](\depscohs[n-1][0,...,p-1])(l),            \\
              \restrp[n][p][q][\epsilon]
              \left(\begin{array}{l}
                  \depscohs[n][0,...,n] \\
                  E                     \\
                  E'                    \\
                \end{array}\right)(\restrf[n][p](\depscohs[n][0,...,p])(d))(c) \\
            \end{array}\right)                                                          \\
  \end{array}
\end{equation*}

@@

              \depscohs[n][0,...,n]:\depscohstype[n][0,...,n] \\
              \cohp[n][0,...,p-2]:\restrptype[n][0,...,p-1]

@@

                  \depscohs[n][0,...,n] \\

@@

              \depscohs[n][0,...,n] \\

@@

                  \depscohs[n+1][0,...,n] \\

@@

              \depscohs[n][0,...,n]:\depscohstype[n][0,...,n]                      \\
              \cohp[n][0,...,p-1]:\cohptype[n][0,...,p-1](\depscohs[n][0,...,p-1]) \\

@@

@@

              \depscohs.\cohf[n][p][r][p][\omega][\theta] \circ \cohf[n+1][p][q+1][p][\epsilon][\theta]

@@

              \circ \depscohs.\cohf[n][p][q][r][\epsilon][\omega] =              \\

@@

              \circ \depscohs.\cohf[n][p][q][p][\epsilon][\theta]

@@

              \depscohs[n][0,...,n]:\depscohstype[n][0,...,n] \\

@@

\begin{align*}
  \depscohttype[n][0,..,p-1] \defeq
  \left(\begin{array}{l}
            \depscohs[n][0,...,n] :\depscohstype[n][0,...,n] \\
            \cohp[n][0,...p-1]: \cohptype[n][0,...,p-1]      \\
            \coht[n][0,...,p-1]: \cohttype[n][0,...,p-1]     \\
          \end{array}\right)
\end{align*}

>>>>

\newcommandx{\deps}[2][1,2]{\prim{deps}[][#1][#2][][]}
\newcommandx{\fulldeps}[2][1,2]{\prim{fulldeps}[][#1][][][]}
\newcommandx{\depscohs}[2][1,2]{\prim{depscohs}[][#1][#2][][]}
\newcommandx{\fulldepscohs}[2][1,2]{\prim{fulldepscohs}[][#1][][][]}
\newcommandx{\depscoht}[2][1,2]{\prim{depscoht}[][#1][#2][][]}
\newcommandx{\fulldepscoht}[2][1,2]{\prim{fulldepscoht}[][#1][][][]}
\newcommandx{\depstype}[2][1,2]{\prim{depstype}[][#1][#2][][]}
\newcommandx{\fulldepstype}[2][1,2]{\prim{fulldepstype}[][#1][][][]}
\newcommandx{\depscohstype}[2][1,2]{\prim{depscohstype}[][#1][#2][][]}
\newcommandx{\fulldepscohstype}[2][1,2]{\prim{fulldepscohstype}[][#1][][][]}
\newcommandx{\depscohttype}[2][1,2]{\prim{depscohttype}[][#1][#2][][]}
\newcommandx{\fulldepscohttype}[2][1,2]{\prim{fulldepscohttype}[][#1][][][]}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n][0,...,p]   &
    (\fulldeps[n-1])        & :      & \paintingtype[n][0,...,p](\framep[n][0,...,p](\fulldeps.\deps[n-1][0,...,p])) \\
    \painting[n][0,...,0-1] &
    \left(\begin{array}{l}
              \unitpoint \\
              E          \\
            \end{array}\right)  & \defeq & \unittype                                                                   \\
    \painting[n][0,...,p]   &
    (\fulldeps[n-1])
                            & \defeq &
    (\painting[n][0,...,p-1](\fulldeps[n-1]),\painting[n][p](\fulldeps[n-1]))                                        \\
  \end{array}
\end{equation*}

@@

\begin{align*}
  \depscohstype[n][0,..,p-1] \defeq
  \left(\begin{array}{l}
            \deps[n][0,...,n] :\depstype[n][0,...,n]        \\
            \restrp[n][0,...p-1]: \restrptype[n][0,...,p-1] \\
            \cohf[n][0,...,p-1]: \cohftype[n][0,...,p-1]    \\
          \end{array}\right)
\end{align*}
so that the type of $\restrf[n][0,...,p]$ can be rephrased as $\Pi \depscohs[n][0,...,p-1]:\depscohstype[n][0,..,p-1]. \restrftype[n][0,...,p]      \left(\begin{array}{l}
      \depscohs.\fulldeps.\deps.\framep[n][0,...,p]   \\
      \depscohs.\fulldeps.\deps.\painting[n][0,...,p] \\
    \end{array}\right)$, and similarly for $\restrltype$ and $\restrl$. Later, we shall write $\depscohs.\deps$, $\depscohs.\restrp$, and $\depscohs.\cohf$ for the corresponding projections of an element $\depscohs$ in $\depscohstype$. By abuse of notation, we also write $\unitpoint$ for the only inhabitant of $\depscohstype[n][0,..,0-1]$.

@@

\begin{align*}
  \overline{\deps}(\depscohs:\depscohstype) & :      & \depstype \\
  \overline{\deps}(\depscohs)               & \defeq &
  \left(\begin{array}{l}
            \framep[n][p](\depscohs.\deps[n][0,...,p])   \\
            \painting[n][p](\depscohs.\deps[n][0,...,n]) \\
            \restrf[n][p](\depscohs[n][0,...,p])         \\
          \end{array}\right)             \\
\end{align*}

We need also the notion of full dependencies at the level of coherence conditions, that combines a full $\depscohs[n][0,...,n]$ with a set $E:\framep[n][n](\overline{\deps}(\depscohs[n][0,...,n-1])) \rightarrow \U$:
\begin{align*}
  \fulldepscohstype[n] \defeq
  \left(\begin{array}{l}
            \depscohs[n][0,...,n-1] :\depscohstype[n][0,...,n-1]                         \\
            E:\framep[n-1][n-1](\overline{\deps}(\depscohs[n][0,...,n-1]))\rightarrow \U \\
          \end{array}\right)
\end{align*}
TODO: say the convention on projection in the Introduction

Also, when $n$ is 0, we write $(E',E)$ to denote an element $(E', E)$ of $\fulldepscohstype[0]$.

Armed with $\restrf$, $\cohftype$ and these abbreviations, it is now possible to define $\restrp[n][p]$ by recursion on $n-p$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrp[n][p]                           &
    (\fulldepscohs[n]:\fulldepscohstype[n]) & :      &
    \left(
    \begin{array}{l}
        \Pi q\leq n. \Pi \epsilon. \Pi d:\framep[n][p]
        \left(\begin{array}{l}
                  \overline{\deps}(\depscohs)    \\
                  E':\framep[n][n]\rightarrow \U \\
                \end{array}\right).    \\
        \painting[n][p]
        \left(\begin{array}{l}
                  \overline{\deps}(\depscohs)    \\
                  E':\framep[n][n]\rightarrow \U \\
                \end{array}\right)(d) \rightarrow \\
        \painting[n-1][p]
        \left(\begin{array}{l}
                  \depscohs[n][0,...,n].\deps       \\
                  E:\framep[n-1][n-1]\rightarrow \U \\
                \end{array}\right)(\restrf[n][p](\depscohs[n][0,...,n].deps)(d))\end{array}\right)                                       \\
    \restrp[n][n]                           &
    (\fulldepscohs[n][0,...,n])             & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c). l\epsilon \\
    \restrp[n][p](p<n)                      &
    (\fulldepscohs[n][0,...,n])             & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c).
    \left(\begin{array}{l}
              \restrl[n][p][q][\epsilon](\fulldepscohs.\depscohs[n-1][0,...,p-1])(l),               \\
              \restrp[n][p][q][\epsilon]
              (\fulldepscohs[n][0,...,n])(\restrf[n][p](\fulldepscohs.\depscohs[n][0,...,p])(d))(c) \\
            \end{array}\right)                              \\
  \end{array}
\end{equation*}

@@

              \fulldepscohs[n]:\fulldepscohstype[n]                                    \\
              \cohp[n][0,...,p-2]:\restrptype[n][0,...,p-1](\fulldepscohs[n][0,...,n]) \\

@@

                  \fulldepscohs[n]    \\
                  \cohp[n][0,...,p-1] \\

@@

              \fulldepscohs[n]  \\

@@

                  \fulldepscohs[n+1]    \\

@@

              \fulldepscohs[n]:\fulldepscohstype[n]                         \\
              \cohp[n][0,...,p-1]:\cohptype[n][0,...,p-1](\fulldepscohs[n]) \\

@@

              (E',E)     \\

@@

              \fulldepscohs.\depscohs.\cohf[n][p][r][p][\omega][\theta] \circ \cohf[n+1][p][q+1][p][\epsilon][\theta]

@@

              \circ \fulldepscohs.\depscohs.\cohf[n][p][q][r][\epsilon][\omega] = \\

@@

              \circ \fulldepscohs.\depscohs.\cohf[n][p][q][p][\epsilon][\theta]

@@

              \fulldepscohs[n]:\fulldepscohstype[n]   \\

@@

\begin{align*}
  \depscohttype[n][0,..,p-1] \defeq
  \left(\begin{array}{l}
            \fulldepscohs[n] :\fulldepscohstype[n]                        \\
            \cohp[n][0,...p-1]: \cohptype[n][0,...,p-1](\fulldepscohs[n]) \\
            \coht[n][0,...,p-1]: \cohttype[n][0,...,p-1]
            \left(\begin{array}{l}
                \fulldepscohs[n]    \\
                \cohp[n][0,...,p-1] \\
              \end{array}\right)                                     \\
          \end{array}\right)
\end{align*}

We need also the notion of full dependencies at the level of the coherence conditions of dimension 2, that combines a full $\depscoht[n][0,...,n]$ with a set $E:\framep[n][n](\overline{\deps}(\overline{\deps}(\depscohs[n][0,...,n-1])))??? \rightarrow \U$:
\begin{align*}
  \fulldepscohttype[n] \defeq
  \left(\begin{array}{l}
            \depscoht[n][0,...,n-1] :\depscohttype[n][0,...,n-1]    \\
            E:\framep[n-1][n-1](\overline{\deps}(??))\rightarrow \U \\
          \end{array}\right)
\end{align*}
